### 2018-04-30 Maths Notes/Mechanics/M2/Centres of mass/Centres of mass.tex
Notes Update

\begin{document}
\begin{center}
\underline{\huge Centres of mass}
\end{center}
\begin{obeylines}
\section{Centre of mass of a discrete mass distribution}
$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$\\
$\bar{y}=\dfrac{\Sigma m_iy_i}{\Sigma m_i}$\\
\textbf{Example}\\
\end{obeylines}
\begin{tabular}{|c|c|c|c|}
 \hline
 Mass&$2$&$3$&$2$\\
 \hline
 $x$&2&3&-3\\
 \hline
 $y$&3&6&2\\
 \hline

\end{tabular}
\\
\\
\begin{obeylines}
$\bar{x}=\dfrac{2\times 2+3\times3+2\times-3}{2+3+2}=1$
$\bar{y}=\dfrac{2\times3+3\times6+2\times2}{2+3+2}=4$
\section{Uniform laminae}
For a triangular lamina the centre of mass is $\frac{2}{3}$ along the line from the vertex to the middle of the line opposite.
For a sector of a circle, radius r, where the angle at the centre is $2\alpha$ the centre of mass is $\dfrac{2r\sin\alpha}{3\alpha}$
\section{Rods}
In a circular arc, radius r, where the angle at the centre is $2\alpha$, the centre of mass is $\dfrac{r\sin\alpha}{\alpha}$ away from the centre.
\section{Equilibrium}
To avoid tipping, the line of action of the weight must be within the side of the lamina in contact with the plane.
If a lamina is suspended from a fixed point, the centre of mass will be vertically below the point of suspension.
Assumptions made in equilibrium calculations:
\begin{itemize}
\item No friction at the point of suspension
\item The mass of each area is uniform
\item The mass is uniform at the join
\end{itemize}

\end{obeylines}
\newpage
\begin{center}
\underline{\huge Centres of mass example - Equilibrium problems}
\end{center}
\includegraphics[width=1.5in]{eq.jpg}\\
\\
\textit{(a) Find the distance of the centre of mass of the lamina from AF. }\\

>>>>

\begin{document}
\begin{center}
\underline{\huge Centres of mass}
\end{center}
\section{Centre of mass of a discrete mass distribution}
$$\bar{x}=\dfrac{\Sigma m_ix_i}{\Sigma m_i}$$
$$\bar{y}=\dfrac{\Sigma m_iy_i}{\Sigma m_i}$$
\subsection{Example}
\begin{tabular}{|c|c|c|c|}
 \hline
 Mass&$2$&$3$&$2$\\
 \hline
 $x$&2&3&-3\\
 \hline
 $y$&3&6&2\\
 \hline

\end{tabular}
\\
$$\bar{x}=\dfrac{2\times 2+3\times3+2\times-3}{2+3+2}=1$$
$$\bar{y}=\dfrac{2\times3+3\times6+2\times2}{2+3+2}=4$$
\section{Uniform laminae}
For a triangular lamina the centre of mass is $\frac{2}{3}$ along the line from the vertex to the middle of the line opposite.\\
\\
If given the coordinates of the vertices of a triangle, the centre of mass is the has coordinates of the mean of the x coordinates, and the mean of the y coordinates.\\
\\
For a sector of a circle, radius r, where the angle at the centre is $2\alpha$ the centre of mass is $\dfrac{2r\sin\alpha}{3\alpha}$
\section{Rods}
In a circular arc, radius r, where the angle at the centre is $2\alpha$, the centre of mass is $\dfrac{r\sin\alpha}{\alpha}$ away from the centre.
\section{Equilibrium}
To avoid tipping, the line of action of the weight must be within the side of the lamina in contact with the plane.
If a lamina is suspended from a fixed point, the centre of mass will be vertically below the point of suspension.
Assumptions made in equilibrium calculations:
\begin{itemize}
\item No friction at the point of suspension
\item The mass of each area is uniform
\item The mass is uniform at the join
\end{itemize}

\newpage
\begin{center}
\underline{\huge Centres of mass example - Equilibrium problems}
\end{center}
\includegraphics[width=1.5in]{eq.jpg}\\
\\
\textit{(a) Find the distance of the centre of mass of the lamina from AF. }\\

### 2018-04-30 Maths Notes/Mechanics/M2/Collisions/Collisions.tex
Update of all files to shorten preamble
Tidied to work with all linux systems
Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}

\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}
\pretolerance=10000
\hyphenpenalty=10000
\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}

@@

$e=\dfrac{\text{Speed of seperation}}{\text{Speed of approach}}$

@@

$mgh=\frac{1}{2}mv^2$\\
$v=\sqrt{2gh}$\\
$e=\dfrac{\sqrt{2gh_2}}{\sqrt{2gh_1}}$\\
$e=\dfrac{\sqrt{h_2}}{\sqrt{h_1}}$\\

@@

When doing calculations involving the coefficient of restitution both the calculation for CoR and conservation of momentum will be needed.\\
\textbf{Conservation of momentum:}
$m_1u_1+m_2u_2=m_1v_1+m_2v_2$\\
\textbf{Coefficient of restitution}\\
$e=\dfrac{v_1}{u_1}$

>>>>

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - M2}

@@

$$e=\frac{\text{Speed of seperation}}{\text{Speed of approach}}$$

@@

$$mgh=\frac{1}{2}mv^2$$
$$v=\sqrt{2gh}$$
$$e=\dfrac{\sqrt{2gh_2}}{\sqrt{2gh_1}}$$
$$e=\dfrac{\sqrt{h_2}}{\sqrt{h_1}}$$

@@

When doing calculations involving the coefficient of restitution both the calculation for CoR and conservation of momentum will be needed.

### 2018-04-30 Maths Notes/Mechanics/M2/Kinematics/Kinematics.tex
Update of all files to shorten preamble
Tidied to work with all linux systems
Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usepackage{tikz}
\usetikzlibrary{positioning}
\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}
\usepackage{mathtools}
\newtagform{brackets}{[}{]}
\usetagform{brackets}
\begin{document}
\begin{center}
\underline{\huge Kinematics}
\end{center}
\begin{obeylines}
\section{Horizontal Projections}
For a \textbf{constant speed} use \textbf{$Speed=\dfrac{Distance}{Time}$}

For a \textbf{constant acceleration} use \textbf{SUVAT}

For all projections:
\begin{itemize}
\item Assume air resistance to be zero
\item Resolve horizontal and vertical motion
\item Horizontal - Constant speed
\item Vertical - Constant acceleration

\section{Angular projections}
The same as horizontal projections but the initial vertical velocity isn't zero.
Example:
A particle is projected at a speed of $49ms^{-1}$ at an angle of $45 \degree$ above the horizontal.
\textit{What is the time taken for the particle to reach its maximum height?}
\item u=$49\sin45$
\item v=0
\item a=-g
\item t=?
$0=49\sin45-gt$
$t=\dfrac{49\sin45}{g}=\dfrac{5\sqrt{2}}{2}\approx3.54$
\textit{What is the maximum height reached?}
\item u=$49\sin45$
\item v=0
\item a=-g
\item s=?
$v^2=u^2+2as$
0=$(49\sin45)^2-2gs$
$S=\dfrac{(49\sin45)^2}{2g}=61.3$
\textit{What is the time of the flight?}
\item u=$49\sin45$
\item a=-g
\item S=0
\item t=?
$S=ut+\dfrac{1}{2}at^2$
$ $
$0=(49\sin45)t-\dfrac{1}{2}gt^2$
0=$t(49\sin45-\dfrac{gt}{2}$
$t=0$
$49\sin45=\dfrac{gt}{2}$
$t=\dfrac{2\times49\sin45}{g}=7.07$
\textit{What is the horizontal range of the particle?}
\item t=7.07
\item Speed=$49\cos45$
$S=45\cos45\times7.07=245$
\end{obeylines}

@@

\textit{At $t = 0$, P is at the origin O. Find the displacement of P from O when}\\

@@

$\mathit{t = 6}$\\
\textbf{Integrate velocity}\\
\textcolor{red}{$\mathlarger{x=\int v \ dt}$}\\
$x=\int 10t-2t^2 \ dt= 5t^2-\frac{2}{3}t^3+c$\\
\textbf{Substitute in the value of t}\\
$t=6$\\
$x=5\times6^2-\frac{2}{3}\times6^3=36m$\\
$\mathit{t = 10}$
\textbf{Integrate velocity for the second half of the journey}\\
$x=\mathlarger{\int-\dfrac{432}{t^2} \ dt=\int-432t^{-2} \ dt=\dfrac{-432t^{-1}}{-1}+k=\dfrac{432}{t}+k}$\\
\textbf{Find value of k by using known distance at t=6}\\
$t=6 \ \ x=36$\\
$36=\dfrac{432}{6}+k$\\
$k=36-72=-36$\\
\textbf{Find distance using value of k and t}\\
$t=10$\\
$x=\dfrac{432}{10}-36=7.2\text{m}$

>>>>

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - M2}

%File specific Preamble 
\usetikzlibrary{positioning} %Used for diagram at bottom

\begin{document}
\begin{center}
\underline{\huge Kinematics}
\end{center}
\section{Horizontal Projections}
For a \textbf{constant speed} use \textbf{$Speed=\dfrac{Distance}{Time}$}\\
For a \textbf{constant acceleration} use \textbf{SUVAT}\\
For all projections:
\begin{itemize}
\item Assume air resistance to be zero
\item Resolve horizontal and vertical motion
\item Horizontal - Constant speed
\item Vertical - Constant acceleration
\end{itemize}
\section{Angular projections}
The same as horizontal projections but the initial vertical velocity isn't zero.
\subsection{Example}
\textit{A particle is projected at a speed of $49ms^{-1}$ at an angle of $45 \degree$ above the horizontal.}\\
\textit{What is the time taken for the particle to reach its maximum height?}

u=$49\sin45$\\
v=0\\
a=-g\\
t=?
$$0=49\sin45-gt$$
$$t=\dfrac{49\sin45}{g}=\dfrac{5\sqrt{2}}{2}\approx3.54$$
\textit{What is the maximum height reached?}\\
u=$49\sin45$\\
v=0\\
a=-g\\
s=?
$$v^2=u^2+2as$$
$$0=(49\sin45)^2-2gs$$
$$S=\dfrac{(49\sin45)^2}{2g}=61.3$$
\textit{What is the time of the flight?}\\
u= $49\sin45$\\
a=-g\\
S=0\\
t=?

$$S=ut+\dfrac{1}{2}at^2$$
$$0=(49\sin45)t-\dfrac{1}{2}gt^2$$
$$49\sin45=\dfrac{gt}{2}$$
$$t=\dfrac{2\times49\sin45}{g}=7.07$$
\textit{What is the horizontal range of the particle?}\\
t=7.07\\
Speed=$49\cos45$
$$S=45\cos45\times7.07=245$$

@@

\textit{At $t = 0$, P is at the origin O. Find the displacement of P from O when} $\mathit{t = 6}$\\

@@

\textbf{Integrate velocity}
\textcolor{red}{$$\mathlarger{x=\int v \ dt}$$}
$$x=\int 10t-2t^2 \ dt= 5t^2-\frac{2}{3}t^3+c$$
\textbf{Substitute in the value of t}
$$t=6$$
$$x=5\times6^2-\frac{2}{3}\times6^3=36m$$
\textit{Find the displacement from O when t = 10}\\
\textbf{Integrate velocity for the second half of the journey}
$$x=\mathlarger{\int-\dfrac{432}{t^2} \ dt=\int-432t^{-2} \ dt=\dfrac{-432t^{-1}}{-1}+k=\dfrac{432}{t}+k}$$
\textbf{Find value of k by using known distance at t=6}
$$t=6 \ \ x=36$$
$$36=\dfrac{432}{6}+k$$
$$k=36-72=-36$$
\textbf{Find distance using value of k and t}
$$t=10$$
$$x=\dfrac{432}{10}-36=7.2\text{m}$$

### 2018-04-30 Maths Notes/Mechanics/M2/Work, energy and power/Work, energy and power.tex
Update of all files to shorten preamble
Tidied to work with all linux systems
Notes Update

\documentclass{article}[18pt]
\usepackage[utf8]{inputenc}
\usepackage[margin=0.7in]{geometry}
\usepackage{parselines} 
\usepackage{amsmath}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{fancyhdr}
\usepackage{gensymb}
\usetikzlibrary{positioning,scopes,decorations.text}
\pgfplotsset{width=10cm,compat=1.9}

\titlespacing\section{0pt}{14pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\newlength\tindent
\setlength{\tindent}{\parindent}
\setlength{\parindent}{0pt}
\renewcommand{\indent}{\hspace*{\tindent}}

\pagestyle{fancy}
\fancyhf{}
\rhead{Sam Robbins 13SE}
\lhead{A Level Maths - M2}
\rfoot{Page \thepage}

\begin{document}
\begin{center}
\underline{\huge Work, energy and power}
\end{center}
\begin{obeylines}
\section{Work}
$\mathbf{Work(J)=Force(N)\times Distance(m)}$
For work done against  gravity, use vertical distance.
\subsection{Work done against friction}
Work done against friction is:
$$WD=\mu R\times \textrm{Distance}$$
Where the distance is in the direction of motion
\section{Energy}
$\mathbf{E_K=\frac{1}{2}mv^2}$
$ $
$\mathbf{GPE=E_P=mgh}$
\section{Conservation of energy}
A particle's total energy is constant if it is subject only to gravity (i.e. smooth surfaces)
If there are frictional forces to consider then the loss of energy is the work done by friction.
\section{Power}
$\mathbf{Power=Force\times Velocity}$
Usually given in kW
\end{obeylines}
\newpage
\begin{center}
\underline{\huge Work, Energy and Power Example - Conservation of energy}
\end{center}
\textit{A block of mass 10 kg is pulled along a straight horizontal road by a constant horizontal force
of magnitude 70 N in the direction of the road. The block moves in a straight line passing
through two points A and B on the road, where AB = 50 m. The block is modelled as a particle and the road is modelled as a rough plane. The coefficient of friction between the block and the road is $\frac{4}{7}$}\\
\\
\textit{Calculate the work done against friction in moving the block from A to B. }\\
\\
Calculate the force of friction
$$F=\mu R=\frac{4}{7}\times10g=56N$$
Multiply force by distance to find work done
$$50\times56=2800J$$
\\
\textit{The block passes through A with a speed of $2ms^{-1}$.\\
Find the speed of the block at B.}
\\
\\
Subtract the work done against friction from the work done by the pulling force
$$70\times50-2800=700J$$
Write an expression for the change in kinetic energy
$$\frac{1}{2}\times10\times v^2-\frac{1}{2}\times10\times2^2$$
Set the excess energy from the work done equal to the change in kinetic energy and simplify
$$700=5v^2-20$$
$$v=12$$
\newpage
\begin{center}
\underline{\huge Work, Energy and Power Example - Power}
\end{center}
\textit{A lorry of mass 2000 kg is moving down a straight road inclined at angle $\alpha$ to the horizontal,where $\sin\alpha=\frac{1}{25}$. The resistance to motion is modelled as a constant force of magnitude 1600 N. The lorry is moving at a constant speed of $14ms^{-1}$.\\
Find, in kW, the rate at which the lorry's engine is working.}
\\
\\
Draw a diagram to represent the information
\def\iangle{35} % Angle of the inclined plane

>>>>

\documentclass{article}[18pt]
\usepackage{../../../../format}
\lhead{A Level Maths - M2}

%File specific Preamble
\usetikzlibrary{positioning,scopes,decorations.text}

\begin{document}
\begin{center}
\underline{\huge Work, energy and power}
\end{center}
\begin{obeylines}
\section{Work}
$$\mathbf{\textrm{Work(J)=Component of force in the direction of motion(N)}\times \textrm{Distance moved in the same direction(m)}}$$
For work done against  gravity, use vertical distance.
\subsection{Work done against friction}
Work done against friction is:
$$WD=\mu R\times \textrm{Distance}$$
Where the distance is in the direction of motion
\section{Energy}
$$E_K=\frac{1}{2}mv^2$$
$$GPE=E_P=mgh$$
\section{Conservation of energy}
A particle's total energy is constant if it is subject only to gravity (i.e. smooth surfaces)
If there are frictional forces to consider then the loss of energy is the work done by friction.
\section{Power}
$$\mathbf{Power=Force\times Velocity}$$
Usually given in kW
\end{obeylines}
\newpage
\begin{center}
\underline{\huge Work, Energy and Power Example - Conservation of energy}
\end{center}
\textit{A block of mass 10 kg is pulled along a straight horizontal road by a constant horizontal force
of magnitude 70 N in the direction of the road. The block moves in a straight line passing
through two points A and B on the road, where AB = 50 m. The block is modelled as a particle and the road is modelled as a rough plane. The coefficient of friction between the block and the road is $\frac{4}{7}$}\\
\\
\textit{Calculate the work done against friction in moving the block from A to B. }\\
\\
Calculate the force of friction
$$F=\mu R=\frac{4}{7}\times10g=56N$$
Multiply force by distance to find work done
$$50\times56=2800J$$
\\
\textit{The block passes through A with a speed of $2ms^{-1}$.\\
Find the speed of the block at B.}
\\
\\
Subtract the work done against friction from the work done by the pulling force
$$70\times50-2800=700J$$
Write an expression for the change in kinetic energy
$$\frac{1}{2}\times10\times v^2-\frac{1}{2}\times10\times2^2$$
Set the excess energy from the work done equal to the change in kinetic energy and simplify
$$700=5v^2-20$$
$$v=12$$
\newpage
\begin{center}
\underline{\huge Work, Energy and Power Example - Power}
\end{center}
\textit{A lorry of mass 2000 kg is moving down a straight road inclined at angle $\alpha$ to the horizontal,where $\sin\alpha=\frac{1}{25}$. The resistance to motion is modelled as a constant force of magnitude 1600 N. The lorry is moving at a constant speed of $14ms^{-1}$.\\
Find, in kW, the rate at which the lorry's engine is working.}
\\
\\
Draw a diagram to represent the information
\def\iangle{35} % Angle of the inclined plane
